report.rmd
INFO 550 Report

---
title: "INFO 550 Report"
author: "Madison Vignes"
date: `r format(Sys.Date(), '%Y-%m-%d')`"
output: pdf_document
---

Load data into R
```{r, loading-data, echo = FALSE}
genes=read.table("hg18genes.txt", comment="", header=TRUE)
```

Cleaning and Analysis 

```{r, label-of-chunk, echo = FALSE}
# basic data cleaning 
genes1 <- na.omit(genes)
#creating a new gene length variable
geneLen=genes1$txEnd-genes1$txStart
```
When we create a histogram of the gene lengths in Figure \@ref(fig:make-a-plot1), we are able to visuralize the distibution of gene lengths by each chomosome. But, as seen by this plot, a normal distribution is not the best fit for this data, therefore another histogram was created using a log scale to visualize data.  

```{r, make-a-plot1, echo = FALSE, fig.cap="Figure 1. The distribution of gene length"}
# histogram of gene length
hist(geneLen,50, main="Gene lengths", xlab="base pairs")
```
As we see in Figure \@ref(fig:make-a-plot2), the log distibution of gene length more accuratley represents the length of genes on each chromosome.

```{r, make-a-plot2, echo = FALSE, fig.cap="Figure 2. The log distribution of gene length"}
# histogram of log gene length - a clear bimodal distribution.
hist(log(geneLen),50, main="log(Gene lengths)", xlab="base pairs"
```
